Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2014-12-18 content/appendix.tex
spacing

\efficient*
\begin{proof}
\leavevmode
  \begin{enumerate}
    \item The identity function is represented as $\lambda r\,.\,\setpair{\bot_T}{\top_T}$. 
    The top function is represented as $\lambda r\,.\,\setpair{\top_T}{\top_T}$.
    \item Lemmas~\ref{lem:equalComp} and \ref{lem:equalMeet} show that the representation of micro functions is closed under composition and meet.
    \item To show that our representation for micro functions forms a lattice with finite height, let us first show that $\lcc_\rcc:\,\rcc\to2^T$ forms a lattice. Since $T$ is a finite set, $(2^T,\,\subseteq)$ is a finite-height lattice. $\rcc$ is a finite set. Hence, the mapping
    \[
      \rcc\mapsto2^T=\{(r,\,t)\,|\,r\in \rcc,\,t\in 2^T\}=\lcc_\rcc
    \]        
    also forms a finite-height lattice \cite{nielson1999principles}.

@@

\commentout{
\begin{proof}[\textbf{Proof of Lemma~\ref{lem:efficient}}]
  \begin{enumerate}
    \item The identity function is represented as
      \[
        \denote{\id}=\{(r,\,\langle \bot_T,\,\top_T\rangle)\,|\,r\in \rcc\}\,;
      \]
      the top function is represented as
      \[
        \denote{\lambda m\,.\,\topcc}=\{(r,\,\langle \top_T,\,\top_T\rangle)\,|\,r\in \rcc\}\,.
      \]
    \item Equations~\eqref{eq:compclosed} and~\eqref{eq:meetclosed} show that the representation of micro functions is closed under composition and meet.
    \item To show that our representation for micro functions forms a lattice with finite height, let us first show that $\lcc_\rcc:\,\rcc\to2^T$ forms a lattice. Since $T$ is a finite set, $(2^T,\,\subseteq)$ is a finite-height lattice. $\rcc$ is a finite set. Hence, the mapping
    \[
      \rcc\mapsto2^T=\{(r,\,t)\,|\,r\in \rcc,\,t\in 2^T\}=\lcc_\rcc
    \]        
    also forms a finite-height lattice \cite{nielson1999principles}.

    Furthermore, $\lcc_\rcc$ is a finite set. 
    Every element of $\lcc_\rcc$ can be applied to $|\rcc|$ receivers, where each receiver is mapped to a set of types. There are $|\rcc|\cdot2^{|T|}$ different possibilities to form those mappings, so
    \[
      |\lcc_\rcc|=|\rcc|\cdot2^{|T|}.
    \]
    Therefore, $\lcc_\rcc\mapsto \lcc_\rcc$ also forms a finite-height lattice.
    \item All operations can be computed in $O(\rcc\times T)$ time. Note that the $\rcc$ and $T$ sets are an input to the correlated-calls analysis, and the time it takes to compute the meet or composition of micro functions is independent of the representation of the specific operand micro functions.
    \item The space bound is $O(\rcc\times T)$.
  \end{enumerate}
\end{proof}
}

>>>>

\efficient*
\begin{proof}
\leavevmode
  \begin{enumerate}
    \item The identity function is represented as $\lambda r\,.\,\setpair{\bot_T}{\top_T}$. 
    The top function is represented as $\lambda r\,.\,\setpair{\top_T}{\top_T}$.
    \item Lemmas~\ref{lem:equalComp} and \ref{lem:equalMeet} show that the representation of micro functions is closed under composition and meet.
    \item To show that our representation for micro functions forms a lattice with finite height, let us first show that $\lcc_\rcc:\,\rcc\to2^T$ forms a lattice. Since $T$ is a finite set, $(2^T,\,\subseteq)$ is a finite-height lattice. $\rcc$ is a finite set. Hence, the mapping
    \[
      \rcc\mapsto2^T=\{(r,\,t)\,|\,r\in \rcc,\,t\in 2^T\}=\lcc_\rcc
    \]        
    also forms a finite-height lattice~\cite{nielson1999principles}.

@@

### 2014-12-18 content/background.tex
spacing

\subsection{Terminology and Notation}
%We will start by introducing several concepts used by the IFDS and IDE analyses.
%\begin{mdelete}\mtodo{We already say (but less precisely) what a CFG is in the introduction.}
The \textit{control-flow graph} of a procedure is a directed graph whose nodes are instructions, and which contains an edge from $n_1$ to $n_2$ whenever
$n_2$ may execute immediately after $n_1$. A CFG has a distinguished \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$.
Following the presentation of \cite{reps1995precise,sagiv1996precise}, we follow every
call instruction with a no-op instruction, so that every \textit{call node} is immediately
followed by a \textit{return node} in the CFG.
%\end{mdelete}
%and edges represent transfer of control between the instructions during an execution of the program. 
%A control-flow graph has a unique start node, $\startmain$, which is the node corresponding to the program entrypoint.
%An \textit{intraprocedural} path is a path in a control-flow graph whose nodes are in the same procedure. By contrast, an \textit{interprocedural path} is one that contains nodes from different procedures. 
The \textit{control-flow supergraph} of a program contains the CFGs of all of the procedures as
subgraphs. In addition, for each call instruction $c$, the supergraph contains a \textit{call-to-start} edge to the start node of every procedure that
may be called from $c$, and an \textit{end-to-return} edge from the end node of the procedure back to the call instruction.
%\otodo{do we really need to extra "return nodes" that the IFDS/IDE papers add?}
%is a control-flow graph
%in which each procedure $p$ is augmented with an additional \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$, and for each call $c_q$ to a procedure $q$, there is a \textit{call node} \textsf{call}$_{c_q}$ and subsequent \textit{return node} \textsf{return}$_{c_q}$.

>>>>

\subsection{Terminology and Notation}
%We will start by introducing several concepts used by the IFDS and IDE analyses.
%\begin{mdelete}\mtodo{We already say (but less precisely) what a CFG is in the introduction.}
The \textit{control-flow graph} of a procedure is a directed graph whose nodes are instructions, and which contains an edge from $n_1$ to $n_2$ whenever
$n_2$ may execute immediately after $n_1$. A CFG has a distinguished \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$.
Following the presentation of Reps~et~al.~\cite{reps1995precise,sagiv1996precise}, we follow every
call instruction with a no-op instruction, so that every \textit{call node} is immediately
followed by a \textit{return node} in the CFG.
%\end{mdelete}
%and edges represent transfer of control between the instructions during an execution of the program. 
%A control-flow graph has a unique start node, $\startmain$, which is the node corresponding to the program entrypoint.
%An \textit{intraprocedural} path is a path in a control-flow graph whose nodes are in the same procedure. By contrast, an \textit{interprocedural path} is one that contains nodes from different procedures. 
The \textit{control-flow supergraph} of a program contains the CFGs of all of the procedures as
subgraphs. In addition, for each call instruction $c$, the supergraph contains a \textit{call-to-start} edge to the start node of every procedure that
may be called from $c$, and an \textit{end-to-return} edge from the end node of the procedure back to the call instruction.
%\otodo{do we really need to extra "return nodes" that the IFDS/IDE papers add?}
%is a control-flow graph
%in which each procedure $p$ is augmented with an additional \textit{start node} \textsf{start}$_p$ and \textit{end node} \textsf{end}$_p$, and for each call $c_q$ to a procedure $q$, there is a \textit{call node} \textsf{call}$_{c_q}$ and subsequent \textit{return node} \textsf{return}$_{c_q}$.

### 2014-12-18 content/example.tex
spacing

In this section, we illustrate our approach using a small motivating example that
shows how our technique can be used to improve the precision of taint analysis.
A taint analysis is a specialized dataflow analysis for computing how string values
may flow from ``sources'', which are typically statements that read untrusted input, 
to ``sinks'', which are typically security-sensitive
operations such as calls to 
a database. In previous research \cite{DBLP:conf/issta/GuarnieriPTDTB11,DBLP:conf/pldi/ArztRFBBKTOM14}, 
taint analysis algorithms have been formulated as IFDS problems.

>>>>

In this section, we illustrate our approach using a small motivating example that
shows how our technique can be used to improve the precision of taint analysis.
A taint analysis is a specialized dataflow analysis for computing how string values
may flow from ``sources'', which are typically statements that read untrusted input, 
to ``sinks'', which are typically security-sensitive
operations such as calls to 
a database. In previous research~\cite{DBLP:conf/issta/GuarnieriPTDTB11,DBLP:conf/pldi/ArztRFBBKTOM14}, 
taint analysis algorithms have been formulated as IFDS problems.

### 2014-12-18 content/intro.tex
spacing

Since a CFG is an over-approximation of the possible flows 
of control in concrete executions of a program, it may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends critically on its ability to detect and disregard such infeasible paths.
The popular \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm 
by  Reps, Horwitz, and Sagiv \cite{reps1995precise} is a general data-flow analysis
algorithm for computing solutions to standard finite distributive data-flow problems such 
as reaching definitions, available expressions, and taint analysis.
A distinguishing 
characteristic of IFDS is that it avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. Sagiv, Reps, and Horwitz
also presented the \textit{Interprocedural Distributive Environment} (IDE) algorithm
\cite{sagiv1996precise} that similarly only considers properly matched call/return
edges, but that supports a broader range of dataflow problems by expanding the 
domain of flow functions to \textit{environments} that go beyond the data-flow facts 
considered by IFDS.

This paper presents an approach to dataflow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
In such cases, if the
method calls are polymorphic (i.e., if they dispatch to different method definitions 
depending on the type of the receiver expression at run time), then their dispatch 
behaviours will be correlated. A recent paper \cite{DBLP:journals/scp/Tip15} identified
this problem but did not present a concrete solution or algorithm, and we are not aware 
of any existing dataflow analysis that is capable of avoiding infeasible paths that arise 
in the presence of correlated method calls.

>>>>

Since a CFG is an over-approximation of the possible flows 
of control in concrete executions of a program, it may contain \textit{infeasible} 
paths that cannot occur at runtime. The precision of a data-flow analysis algorithm
depends critically on its ability to detect and disregard such infeasible paths.
The popular \textit{Interprocedural Finite Distributive Subset} (IFDS) algorithm 
by  Reps, Horwitz, and Sagiv~\cite{reps1995precise} is a general data-flow analysis
algorithm for computing solutions to standard finite distributive data-flow problems such 
as reaching definitions, available expressions, and taint analysis.
A distinguishing 
characteristic of IFDS is that it avoids infeasible interprocedural paths in which calls 
and returns to/from functions are not properly matched. Sagiv, Reps, and Horwitz
also presented the \textit{Interprocedural Distributive Environment} (IDE) algorithm
\cite{sagiv1996precise} that similarly only considers properly matched call/return
edges, but that supports a broader range of dataflow problems by expanding the 
domain of flow functions to \textit{environments} that go beyond the data-flow facts 
considered by IFDS.

This paper presents an approach to dataflow analysis that avoids a type
of infeasible path that arises in object-oriented programs when two or more
methods are dynamically dispatched on the same receiver object.
%\mtodo{Should we define the term ``receiver''?}
In such cases, if the
method calls are polymorphic (i.e., if they dispatch to different method definitions 
depending on the type of the receiver expression at run time), then their dispatch 
behaviours will be correlated. A recent paper~\cite{DBLP:journals/scp/Tip15} identified
this problem but did not present a concrete solution or algorithm, and we are not aware 
of any existing dataflow analysis that is capable of avoiding infeasible paths that arise 
in the presence of correlated method calls.
